Case Study: 16S microbiome data (Schubert et al., 2014)

### 2018-05-13 datasets/RNA-Seq/RNAseq-mir200c.Rmd
Hyperlink DOIs against preferred resolver

As a second RNA-seq dataset, we will test for differences in gene expression upon the knockout of the microRNA mir-200c [(Kim et al., 2013)](http://dx.doi.org/10.1038/nsmb.2701).  The raw fastq files can be found under the accession number `SRP030475`. As the number of samples is limited the experiment might be underpowered, as in most RNA-seq analysis. This is an experimental scenario that could benefit from power gained using modern FDR control methods.

@@

## project data/results folders
datdir <- "data"
resdir <- "results"
dir.create(datdir, showWarnings = FALSE)
dir.create(resdir, showWarnings = FALSE)

## intermediary files we create below
count_file <- file.path(datdir, "rse_gene.Rdata")
result_file <- file.path(resdir, "mir200c-results.rds")
bench_file <- file.path(resdir, "mir200c-benchmark.rds")

@@

@@

## Multiple-Testing Correction

We use the common `BenchDesign` with the set of multiple testing correction methods already included.

```{r}
if (file.exists(bench_file)) {
    sb <- readRDS(bench_file)
} else {
    bd <- initializeBenchDesign()
    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
    saveRDS(sb, file = bench_file)
}
```

There are some warnings from both `BH` and `scott-empirical`. However, they do return results.
I tried to increase the `nmids` parameter for Scott's FDR Regression but this does not appear to make a difference.

@@

## Benchmark Metrics

`scott-empirical` rejects many more hypothesis than the rest of the methods, followed by `lfdr` and `ihw`.

@@

>>>>

As a second RNA-seq dataset, we will test for differences in gene expression upon the knockout of the microRNA mir-200c [(Kim et al., 2013)](https://doi.org/10.1038/nsmb.2701).  The raw fastq files can be found under the accession number `SRP030475`. As the number of samples is limited the experiment might be underpowered, as in most RNA-seq analysis. This is an experimental scenario that could benefit from power gained using modern FDR control methods.

@@

## project data/results folders
datdir <- "data"
resdir <- "results"
sbdir <- "../../results/RNAseq"
dir.create(datdir, showWarnings = FALSE)
dir.create(resdir, showWarnings = FALSE)
dir.create(sbdir, showWarnings = FALSE)

## intermediary files we create below
count_file <- file.path(datdir, "rse_gene.Rdata")
result_file <- file.path(resdir, "mir200c-results.rds")
bench_file <- file.path(sbdir, "mir200c-benchmark.rds")
bench_file_uninf <- file.path(sbdir, "mir200c-uninf-benchmark.rds")

@@

Add random (uninformative) covariate.

```{r}
set.seed(4719)
res$rand_covar <- rnorm(nrow(res))
```

@@

### Random

```{r, rand-diag-scatter, fig.width=4.5, fig.height=3.5}
rank_scatter(res, pvalue = "pval", covariate = "rand_covar") +
    ggtitle("Random independent covariate")
```

```{r, rand-diag-hist, fig.width=10, fig.height=3.2}
strat_hist(res, pvalue = "pval",
           covariate = "rand_covar", maxy = 7.5)
```

## Multiple-Testing Correction - mean

We use the common `BenchDesign` with the set of multiple testing correction 
methods already included. We also add in Scott's FDR Regression (both
`nulltype = "empirical"` and `nulltype = "theoretical"`)
since our test statistics are t-distributed.

```{r}
if (file.exists(bench_file)) {
    sb <- readRDS(bench_file)
} else {
    bd <- initializeBenchDesign()

    bd <- addBMethod(bd, "fdrreg-t",
                     FDRreg::FDRreg,
                     function(x) { x$FDR },
                     z = test_statistic,
                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
                     nulltype = 'theoretical',
                     control = list(lambda = 0.01))

   bd <- addBMethod(bd, "fdrreg-e",
                     FDRreg::FDRreg,
                     function(x) { x$FDR },
                     z = test_statistic,
                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
                     nulltype = 'empirical',
                     control = list(lambda = 0.01))

    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
    saveRDS(sb, file = bench_file)
}
```

There are some warnings from both `BH` and `fdrreg-empirical`. However, they do return results.
I tried to increase the `nmids` parameter for Scott's FDR Regression but this does not appear to make a difference.

@@

## Benchmark Metrics - mean

`fdrreg-empirical` rejects many more hypothesis than the rest of the methods, followed by `lfdr` and `ihw`.

@@

## Multiple-Testing Correction - random

We use the common `BenchDesign` with the set of multiple testing correction 
methods already included. We also add in Scott's FDR Regression (both
`nulltype = "empirical"` and `nulltype = "theoretical"`)
since our test statistics are t-distributed.

```{r}
if (file.exists(bench_file_uninf)) {
    sb <- readRDS(bench_file_uninf)
} else {
    bd <- initializeBenchDesign()

    bd <- addBMethod(bd, "fdrreg-t",
                     FDRreg::FDRreg,
                     function(x) { x$FDR },
                     z = test_statistic,
                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
                     nulltype = 'theoretical',
                     control = list(lambda = 0.01))

   bd <- addBMethod(bd, "fdrreg-e",
                     FDRreg::FDRreg,
                     function(x) { x$FDR },
                     z = test_statistic,
                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
                     nulltype = 'empirical',
                     control = list(lambda = 0.01))

    res <- res %>% dplyr::mutate(ind_covariate = rand_covar)
    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
    saveRDS(sb, file = bench_file_uninf)
}
```

## Benchmark Metrics - random

# Covariate comparison

Here we compare the method ranks for the different covariates at alpha = 0.10.

```{r}
plotMethodRanks(c(bench_file, bench_file_uninf), 
                colLabels = c("mean", "uninf"), 
                alpha = 0.10, xlab = "Comparison")
```

### 2018-05-13 datasets/microbiome/microbiome-baxter.Rmd
Hyperlink DOIs against preferred resolver

Here we download and analyze the Baxter colorectal cancer (CRC) dataset
[(Baxter et al., 2016)](http://dx.doi.org/10.1186/s13073-016-0290-3).
The dataset includes stool samples from 172 patients with no colonic lesions,
and 120 patients with CRC. We'll download the processed OTU tables from
Zenodo and unzip them in the `benchmark-fdr/datasets/microbiome/data` folder.

@@

gns_result_file <- file.path(resdir, "baxter-genus-results.rds")
gns_bench_file <- file.path(sbdir, "baxter-genus-benchmark.rds")
gns_bench_file_log_ubiquity <- file.path(sbdir, "baxter-genus-log-ubiquity-benchmark.rds")
```

@@

>>>>

Here we download and analyze the Baxter colorectal cancer (CRC) dataset
[(Baxter et al., 2016)](https://doi.org/10.1186/s13073-016-0290-3).
The dataset includes stool samples from 172 patients with no colonic lesions,
and 120 patients with CRC. We'll download the processed OTU tables from
Zenodo and unzip them in the `benchmark-fdr/datasets/microbiome/data` folder.

@@

gns_result_file <- file.path(resdir, "baxter-genus-results.rds")
gns_bench_file <- file.path(sbdir, "baxter-genus-benchmark.rds")
gns_bench_file_log_ubiquity <- file.path(sbdir, "baxter-genus-log-ubiquity-benchmark.rds")
gns_bench_file_abun <- file.path(sbdir, "baxter-genus-mean-abun-benchmark.rds")
```

@@

## Multiple-Testing Correction - mean abundance

Let's use `mean_abun_present` as our `ind_covariate`.

```{r, baxter-genus-labelcovariate-abun}
res <- dplyr::mutate(res, ind_covariate = mean_abun_present)
```

```{r, baxter-genus-sb-abun, results="hide", message=FALSE, eval=TRUE}
if (!file.exists(gns_bench_file_abun)) {
    bd <- initializeBenchDesign()
    bd <- dropBMethod(bd, "ashq")
    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
    metadata(sb)$data_download_link <-
                   "https://zenodo.org/record/840333/files/crc_baxter_results.tar.gz"
    saveRDS(sb, file = gns_bench_file_abun)
} else {
    sb <- readRDS(gns_bench_file_abun)
}
```
## Benchmark Metrics - mean abundance

```{r, baxter-genus-abun-metrics, eval=TRUE}
assayNames(sb) <- "qvalue"
sb <- addDefaultMetrics(sb)
```

```{r, baxter-genus-abun-plot, width=15, height=15, eval=TRUE}
rejections_scatter(sb, as_fraction=FALSE, supplementary=FALSE)
rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
```

```{r, baxter-genus-abun-overlap}
plotFDRMethodsOverlap(sb, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
```
```{r}
covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
```

# OTU/Genus comparison

Here we compare the method ranks for the different comparisons at alpha = 0.10.

```{r}
plotMethodRanks(c(otu_bench_file, otu_bench_file_log_ubiquity, 
                  gns_bench_file, gns_bench_file_log_ubiquity,
                  gns_bench_file_abun), 
                colLabels = c("OTU", "OTU - log", "Genus", "Genus - log",
                              "Genus - abun"), 
                alpha = 0.10, xlab = "Comparison")
```

### 2018-05-13 datasets/microbiome/microbiome-enigma.Rmd
Hyperlink DOIs against preferred resolver

Here we analyze a subset of the ENIGMA data ([Smith, M. and Rocha, A. et al 2015](http://dx.doi.org/10.1128/mBio.00326-15)).
The dataset includes water samples from 67 wells with well-characterized biochemical metadata (i.e. pH, heavy metal concentrations, etc).
We acquired the data from a collaborator and have put it in the `benchmark-fdr/datasets/microbiome/DATA/enigma` folder.

@@

# From their paper, we expect SO4 to have little correlation, pH
# to have high correlations, and Al to be in the middle
so4_result_file <- file.path(resdir, "enigma-so4-otus-results.rds")
so4_bench_file <- file.path(sbdir, "enigma-so4-otus-benchmark.rds")

ph_result_file <- file.path(resdir, "enigma-ph-otus-results.rds")
ph_bench_file <- file.path(sbdir, "enigma-ph-otus-benchmark.rds")

al_result_file <- file.path(resdir, "enigma-al-otus-results.rds")
al_bench_file <- file.path(sbdir, "enigma-al-otus-benchmark.rds")
```

@@

## Covariate Diagnostics

@@

### Ubiquity

@@

### Mean Abundance (across non-zero samples)

@@

## Multiple-Testing Correction

@@

## Benchmark Metrics

@@

@@

## Covariate Diagnostics

@@

## Multiple-Testing Correction

@@

@@

```{r, al-pvals}
if (!file.exists(al_result_file)) {
    res <- test_microbiome_corr(abundance = abun_otu, meta_col = meta$Al_mgL,
                                shift = zeroabun)
    saveRDS(res, file = al_result_file)
} else {
    res <- readRDS(al_result_file)
}
```

## Covariate Diagnostics

@@

## Multiple-Testing Correction

@@

>>>>

Here we analyze a subset of the ENIGMA data ([Smith, M. and Rocha, A. et al 2015](https://doi.org/10.1128/mBio.00326-15)).
The dataset includes water samples from 67 wells with well-characterized biochemical metadata (i.e. pH, heavy metal concentrations, etc).
We acquired the data from a collaborator and have put it in the `benchmark-fdr/datasets/microbiome/DATA/enigma` folder.

@@

# From their paper, we expect SO4 to have little correlation, pH
# to have high correlations, and Al to be in the middle
so4_result_file <- file.path(resdir, "enigma-so4-otus-results.rds")
so4_bench_file <- file.path(sbdir, "enigma-so4-otus-benchmark.rds")
so4_bench_file_abun <- file.path(sbdir, "enigma-so4-otus-abun-benchmark.rds")
so4_bench_file_uninf <- file.path(sbdir, "enigma-so4-otus-uninf-benchmark.rds")

ph_result_file <- file.path(resdir, "enigma-ph-otus-results.rds")
ph_bench_file <- file.path(sbdir, "enigma-ph-otus-benchmark.rds")
ph_bench_file_abun <- file.path(sbdir, "enigma-ph-otus-abun-benchmark.rds")
ph_bench_file_uninf <- file.path(sbdir, "enigma-ph-otus-uninf-benchmark.rds")

al_result_file <- file.path(resdir, "enigma-al-otus-results.rds")
al_bench_file <- file.path(sbdir, "enigma-al-otus-benchmark.rds")
al_bench_file_abun <- file.path(sbdir, "enigma-al-otus-abun-benchmark.rds")
al_bench_file_uninf <- file.path(sbdir, "enigma-al-otus-uninf-benchmark.rds")
```

@@

Add random (uninformative) covariate.

```{r}
set.seed(44318)
res$rand_covar <- rnorm(nrow(res))
```

### Covariate Diagnostics

@@

#### Ubiquity

@@

#### Mean Abundance (across non-zero samples)

@@

#### Random

```{r, ph-rand-hist, fig.width=10, fig.height=3.2}
strat_hist(res, pvalue="pval", covariate="rand_covar", maxy=8, binwidth=0.05)
```

```{r, ph-rand-scatter, fig.width=4.5, fig.height=3.5}
rank_scatter(res, pvalue="pval", covariate="rand_covar")
```

### Multiple-Testing Correction - ubiquity

@@

### Benchmark Metrics - ubiquity

@@

### Multiple-Testing Correction - mean abundance

Let's use `mean_abun_present` as our `ind_covariate`.

```{r, ph-labelcovariate-abun}
res <- dplyr::mutate(res, ind_covariate = mean_abun_present)
```

```{r, ph-sb-abun, results="hide", message=FALSE, eval=TRUE}
if (!file.exists(ph_bench_file_abun)) {
    bd <- initializeBenchDesign()
    bd <- dropBMethod(bd, "ashq")
    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
    saveRDS(sb, file = ph_bench_file_abun)
} else {
    sb <- readRDS(ph_bench_file_abun)
}
```
### Benchmark Metrics - mean abundance

```{r, ph-metrics-abun, eval=TRUE}
assayNames(sb) <- "qvalue"
sb <- addDefaultMetrics(sb)
```

```{r, ph-plot-abun, width=15, height=15, eval=TRUE}
rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
```
```{r}
covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
```

### Multiple-Testing Correction - random

Let's use `rand_covar` as our `ind_covariate`.

```{r, ph-labelcovariate-rand}
res <- dplyr::mutate(res, ind_covariate = rand_covar)
```

```{r, ph-sb-rand, results="hide", message=FALSE, eval=TRUE}
if (!file.exists(ph_bench_file_uninf)) {
    bd <- initializeBenchDesign()
    bd <- dropBMethod(bd, "ashq")
    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
    saveRDS(sb, file = ph_bench_file_uninf)
} else {
    sb <- readRDS(ph_bench_file_uninf)
}
```
### Benchmark Metrics - random

```{r, ph-metrics-rand, eval=TRUE}
assayNames(sb) <- "qvalue"
sb <- addDefaultMetrics(sb)
```

```{r, ph-plot-rand, width=15, height=15, eval=TRUE}
rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
```
```{r}
covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
```

@@

Add random (uninformative) covariate.

```{r}
set.seed(27349)
res$rand_covar <- rnorm(nrow(res))
```

### Covariate Diagnostics

@@

#### Random

```{r, so4-rand-hist, fig.width=10, fig.height=3.2}
strat_hist(res, pvalue="pval", covariate="rand_covar", maxy=8, binwidth=0.05)
```

```{r, so4-rand-scatter, fig.width=4.5, fig.height=3.5}
rank_scatter(res, pvalue="pval", covariate="rand_covar")
```

### Multiple-Testing Correction - ubiquity

@@

### Multiple-Testing Correction - mean abundance

Let's use `mean_abun_present` as our `ind_covariate`.

```{r, so4-labelcovariate-abun}
res <- dplyr::mutate(res, ind_covariate = mean_abun_present)
```

```{r, so4-sb-abun, results="hide", message=FALSE, eval=TRUE}
if (!file.exists(so4_bench_file_abun)) {
    bd <- initializeBenchDesign()
    bd <- dropBMethod(bd, "ashq")
    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
    saveRDS(sb, file = so4_bench_file_abun)
} else {
    sb <- readRDS(so4_bench_file_abun)
}
```
### Benchmark Metrics - mean abundance

```{r, so4-metrics-abun, eval=TRUE}
assayNames(sb) <- "qvalue"
sb <- addDefaultMetrics(sb)
```

```{r, so4-plot-abun, width=15, height=15, eval=TRUE}
rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
```
```{r}
covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
```

### Multiple-Testing Correction - random

Let's use `rand_covar` as our `ind_covariate`.

```{r, so4-labelcovariate-rand}
res <- dplyr::mutate(res, ind_covariate = mean_abun_present)
```

```{r, so4-sb-rand, results="hide", message=FALSE, eval=TRUE}
if (!file.exists(so4_bench_file_uninf)) {
    bd <- initializeBenchDesign()
    bd <- dropBMethod(bd, "ashq")
    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
    saveRDS(sb, file = so4_bench_file_uninf)
} else {
    sb <- readRDS(so4_bench_file_uninf)
}
```
### Benchmark Metrics - random

```{r, so4-metrics-rand, eval=TRUE}
assayNames(sb) <- "qvalue"
sb <- addDefaultMetrics(sb)
```

```{r, so4-plot-rand, width=15, height=15, eval=TRUE}
rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
```
```{r}
covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
```

@@

```{r, al-pvals}
if (!file.exists(al_result_file)) {
    res <- test_microbiome_corr(abundance = abun_otu, meta_col = meta$Al_mgL,
                                shift = zeroabun)
    saveRDS(res, file = al_result_file)
} else {
    res <- readRDS(al_result_file)
}

# remove OTUs with missing pvals
res <- res %>% filter(!is.na(pval))
```

Add random (uninformative) covariate.

```{r}
set.seed(19474)
res$rand_covar <- rnorm(nrow(res))
```

### Covariate Diagnostics

@@

#### Random

```{r, al-rand-hist, fig.width=10, fig.height=3.2}
strat_hist(res, pvalue="pval", covariate="rand_covar", maxy=8, binwidth=0.05)
```

```{r, al-rand-scatter, fig.width=4.5, fig.height=3.5}
rank_scatter(res, pvalue="pval", covariate="rand_covar")
```

### Multiple-Testing Correction - ubiquity

@@

### Multiple-Testing Correction - mean abundance

Let's use `mean_abun_present` as our `ind_covariate`.

```{r, al-labelcovariate-abun}
res <- dplyr::mutate(res, ind_covariate = mean_abun_present)
```

```{r, al-sb-abun, results="hide", message=FALSE, eval=TRUE}
if (!file.exists(al_bench_file_abun)) {
    bd <- initializeBenchDesign()
    bd <- dropBMethod(bd, "ashq")
    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
    saveRDS(sb, file = al_bench_file_abun)
} else {
    sb <- readRDS(al_bench_file_abun)
}
```
### Benchmark Metrics - mean abundance

```{r, al-metrics-abun, eval=TRUE}
assayNames(sb) <- "qvalue"
sb <- addDefaultMetrics(sb)
```

```{r, al-plot-abun, width=15, height=15, eval=TRUE}
rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
```
```{r}
covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
```

### Multiple-Testing Correction - random

Let's use `rand_covar` as our `ind_covariate`.

```{r, al-labelcovariate-rand}
res <- dplyr::mutate(res, ind_covariate = rand_covar)
```

```{r, al-sb-rand, results="hide", message=FALSE, eval=TRUE}
if (!file.exists(al_bench_file_uninf)) {
    bd <- initializeBenchDesign()
    bd <- dropBMethod(bd, "ashq")
    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
    saveRDS(sb, file = al_bench_file_uninf)
} else {
    sb <- readRDS(al_bench_file_uninf)
}
```
### Benchmark Metrics - random

```{r, al-metrics-rand, eval=TRUE}
assayNames(sb) <- "qvalue"
sb <- addDefaultMetrics(sb)
```

```{r, al-plot-rand, width=15, height=15, eval=TRUE}
rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
```
```{r}
covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
```

# Response/Covariate comparison

Here we compare the method ranks for the different comparisons at alpha = 0.10.

```{r, width = 15, height = 10}
plotMethodRanks(c(so4_bench_file, so4_bench_file_abun, so4_bench_file_uninf,
                  ph_bench_file, ph_bench_file_abun, ph_bench_file_uninf,
                  al_bench_file, al_bench_file_abun, al_bench_file_uninf), 
                colLabels = c("SO4-ubiquity", "SO4-abun", "SO4-uninf",
                              "pH-ubiquity", "ph-abun", "pH-uninf",
                              "Al-ubiquity", "Al-abun", "Al-uninf"),
                alpha = 0.10, xlab = "Comparison")
```

### 2018-05-13 datasets/microbiome/microbiome-goodrich.Rmd
Hyperlink DOIs against preferred resolver

Here we download and analyze the Goodrich obesity (OB) dataset
[(Goodrich et al., 2014)](http://dx.doi.org/10.1016/j.cell.2014.09.053).
The dataset includes fecal samples from 428 individuals with lean body mass index (BMI < 25),
and 185 obese individuals (BMI > 30). We'll download the processed OTU tables from
Zenodo and unzip them in the `benchmark-fdr/datasets/microbiome/data` folder.

We expect few or no truly differentially abundant OTUs ([Finucane et al., 2014](http://dx.doi.org/10.1371/journal.pone.0084689) and [Walters et al., 2014](https://dx.doi.org/10.1016%2Fj.febslet.2014.09.039)), so this study provides a real-life "null" example.

@@

otu_result_file <- file.path(resdir, "goodrich-otus-results.rds")
otu_bench_file <- file.path(sbdir, "goodrich-otus-benchmark.rds")
gns_result_file <- file.path(resdir, "goodrich-genus-results.rds")
gns_bench_file <- file.path(sbdir, "goodrich-genus-benchmark.rds")
```

@@

## Multiple-Testing Correction

@@

## Benchmark Metrics

@@

@@

## Multiple-Testing Correction

@@

>>>>

Here we download and analyze the Goodrich obesity (OB) dataset
[(Goodrich et al., 2014)](https://doi.org/10.1016/j.cell.2014.09.053).
The dataset includes fecal samples from 428 individuals with lean body mass index (BMI < 25),
and 185 obese individuals (BMI > 30). We'll download the processed OTU tables from
Zenodo and unzip them in the `benchmark-fdr/datasets/microbiome/data` folder.

We expect few or no truly differentially abundant OTUs ([Finucane et al., 2014](https://doi.org/10.1371/journal.pone.0084689) and [Walters et al., 2014](https://doi.org/10.1016%2Fj.febslet.2014.09.039)), so this study provides a real-life "null" example.

@@

otu_result_file <- file.path(resdir, "goodrich-otus-results.rds")
otu_bench_file <- file.path(sbdir, "goodrich-otus-benchmark.rds")
gns_result_file <- file.path(resdir, "goodrich-genus-results.rds")
gns_bench_file <- file.path(sbdir, "goodrich-genus-benchmark.rds")
gns_bench_file_abun <- file.path(sbdir, "goodrich-genus-abun-benchmark.rds")
gns_bench_file_uninf <- file.path(sbdir, "goodrich-genus-uninf-benchmark.rds")
```

@@

## Multiple-Testing Correction - ubiquity

@@

## Benchmark Metrics - ubiquity

@@

Add random (uninformative) covariate.

```{r}
set.seed(76291)
res$rand_covar <- rnorm(nrow(res))
```

@@

### Random (uninformative)

```{r, goodrich-uninf-hist, fig.width=10, fig.height=3.2}
strat_hist(res, pvalue="pval", covariate="rand_covar", maxy=5, binwidth=0.05, numQ=4)
```

```{r, goodrich-uninf-scatter, fig.width=4.5, fig.height=3.5}
rank_scatter(res, pvalue="pval", covariate="rand_covar")
```

## Multiple-Testing Correction - ubiquity

@@

## Multiple-Testing Correction - mean abun

Let's use `mean_abun_present` as our `ind_covariate`.

```{r, goodrich-genus-labelcovariate-abun}
res <- dplyr::mutate(res, ind_covariate = mean_abun_present)
```

```{r, goodrich-genus-sb-abun, results="hide", message=FALSE, eval=TRUE}
if (!file.exists(gns_bench_file_abun)) {
    bd <- initializeBenchDesign()
    bd <- dropBMethod(bd, "ashq")
    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
    metadata(sb)$data_download_link <-
                   "https://zenodo.org/record/840333/files/ob_goodrich_results.tar.gz"
    saveRDS(sb, file = gns_bench_file_abun)
} else {
    sb <- readRDS(gns_bench_file_abun)
}
```
## Benchmark Metrics - mean abun

```{r, goodrich-genus-metrics-abun, eval=TRUE}
assayNames(sb) <- "qvalue"
sb <- addDefaultMetrics(sb)
```

```{r, goodrich-genus-plot-abun, width=15, height=15, eval=TRUE}
rejections_scatter(sb, as_fraction=FALSE, supplementary=FALSE)
rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
```
```{r}
covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
```

```{r, goodrich-genus-overlap-abun}
plotFDRMethodsOverlap(sb, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
```

```{r, goodrich-genus-covariates-abun}
methods <- c( "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf" )
plotCovariateBoxplots(sb, alpha=0.1, nsets=6, methods=methods)
```

## Multiple-Testing Correction - random

Let's use `rand_covar` as our `ind_covariate`.

```{r, goodrich-genus-labelcovariate-uninf}
res <- dplyr::mutate(res, ind_covariate = rand_covar)
```

```{r, goodrich-genus-sb-uninf, results="hide", message=FALSE, eval=TRUE}
if (!file.exists(gns_bench_file_uninf)) {
    bd <- initializeBenchDesign()
    bd <- dropBMethod(bd, "ashq")
    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
    metadata(sb)$data_download_link <-
                   "https://zenodo.org/record/840333/files/ob_goodrich_results.tar.gz"
    saveRDS(sb, file = gns_bench_file_uninf)
} else {
    sb <- readRDS(gns_bench_file_uninf)
}
```
## Benchmark Metrics - random

```{r, goodrich-genus-metrics-uninf, eval=TRUE}
assayNames(sb) <- "qvalue"
sb <- addDefaultMetrics(sb)
```

```{r, goodrich-genus-plot-uninf, width=15, height=15, eval=TRUE}
rejections_scatter(sb, as_fraction=FALSE, supplementary=FALSE)
rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
```
```{r}
covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
```

```{r, goodrich-genus-overlap-uninf}
plotFDRMethodsOverlap(sb, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
```

```{r, goodrich-genus-covariates-uninf}
methods <- c( "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf" )
plotCovariateBoxplots(sb, alpha=0.1, nsets=6, methods=methods)
```

# OTU/Genus comparison

Here we compare the method ranks for the different comparisons at alpha = 0.10.

```{r}
plotMethodRanks(c(otu_bench_file, gns_bench_file, gns_bench_file_abun, gns_bench_file_uninf), 
                colLabels = c("OTU", "Genus-ubiquity", "Genus-abun", "Genus-uninf"), 
                alpha = 0.10, xlab = "Comparison")
```

### 2018-05-13 datasets/microbiome/microbiome-schubert.Rmd
Hyperlink DOIs against preferred resolver

Here we download and analyze the Schubert diarrhea dataset [(Schubert et al., 2014)](http://dx.doi.org/10.1128/mBio.01021-14).
The dataset includes 154 healthy nondiarrheal, 93 Clostridium difficile infection (CDI) associated diarrheal,
and 89 non-CDI associated diarrheal stool samples. We'll download the processed OTU tables from
Zenodo and unzip them in the `benchmark-fdr/datasets/microbiome/data` folder.

@@

otu_result_file <- file.path(resdir, "schubert-otus-results.rds")
otu_bench_file <- file.path(sbdir, "schubert-otus-benchmark.rds")
```

@@

@@

```{r, schubert-abun, fig.width=10, fig.height=3.2, eval=TRUE}
strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=8, binwidth=0.05)
```

@@

## Multiple-Testing Correction

@@

>>>>

Here we download and analyze the Schubert diarrhea dataset [(Schubert et al., 2014)](https://doi.org/10.1128/mBio.01021-14).
The dataset includes 154 healthy nondiarrheal, 93 Clostridium difficile infection (CDI) associated diarrheal,
and 89 non-CDI associated diarrheal stool samples. We'll download the processed OTU tables from
Zenodo and unzip them in the `benchmark-fdr/datasets/microbiome/data` folder.

@@

otu_result_file <- file.path(resdir, "schubert-otus-results.rds")
otu_bench_file <- file.path(sbdir, "schubert-otus-benchmark.rds")
otu_bench_file_abun <- file.path(sbdir, "schubert-otus-abun-benchmark.rds")
otu_bench_file_uninf <- file.path(sbdir, "schubert-otus-uninf-benchmark.rds")
```

@@

Add random (uninformative) covariate.

```{r}
set.seed(9226)
res$rand_covar <- rnorm(nrow(res))
```

@@

```{r, schubert-abun, fig.width=10, fig.height=3.2, eval=TRUE}
strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=17, binwidth=0.05)
```

@@

### Random

```{r, schubert-rand, fig.width=10, fig.height=3.2, eval=TRUE}
strat_hist(res, pvalue="pval", covariate="rand_covar", maxy=17, binwidth=0.05)
```

```{r, schubert-rand-scatter, fig.width=4.5, fig.height=3.5}
rank_scatter(res, pvalue="pval", covariate="rand_covar")
```

## Multiple-Testing Correction - ubiquity

@@

## Multiple-Testing Correction - mean abundance

Let's use `mean_abun_present` as our `ind_covariate`.

```{r, schubert-labelcovariate-abun, eval=TRUE}
res <- dplyr::mutate(res, ind_covariate = mean_abun_present)
```

```{r, schubert-sb-abun, results="hide", message=FALSE, eval=TRUE}
if (!file.exists(otu_bench_file_abun)) {
    bd <- initializeBenchDesign()
    bd <- dropBMethod(bd, "ashq")
    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
    metadata(sb)$data_download_link <-
                   "https://zenodo.org/record/840333/files/cdi_schubert_results.tar.gz"
    saveRDS(sb, file = otu_bench_file_abun)
} else {
    sb <- readRDS(otu_bench_file_abun)
}
```

## Benchmark Metrics (mean abun)

```{r, schubert-metrics-abun, eval=TRUE}
assayNames(sb) <- "qvalue"
sb <- addDefaultMetrics(sb)
```

```{r, schubert-plot-abun, width=15, height=15, eval=TRUE}
rejections_scatter(sb, as_fraction=FALSE, supplementary=FALSE)
rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
plotFDRMethodsOverlap(sb, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
```
```{r}
covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
```

## Multiple-Testing Correction - random

Let's use `rand_covar` as our `ind_covariate`.

```{r, schubert-labelcovariate-rand, eval=TRUE}
res <- dplyr::mutate(res, ind_covariate = rand_covar)
```

```{r, schubert-sb-rand, results="hide", message=FALSE, eval=TRUE}
if (!file.exists(otu_bench_file_uninf)) {
    bd <- initializeBenchDesign()
    bd <- dropBMethod(bd, "ashq")
    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
    metadata(sb)$data_download_link <-
                   "https://zenodo.org/record/840333/files/cdi_schubert_results.tar.gz"
    saveRDS(sb, file = otu_bench_file_uninf)
} else {
    sb <- readRDS(otu_bench_file_uninf)
}
```

## Benchmark Metrics - random

```{r, schubert-metrics-rand, eval=TRUE}
assayNames(sb) <- "qvalue"
sb <- addDefaultMetrics(sb)
```

```{r, schubert-plot-rand, width=15, height=15, eval=TRUE}
rejections_scatter(sb, as_fraction=FALSE, supplementary=FALSE)
rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
plotFDRMethodsOverlap(sb, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
```
```{r}
covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
```

# Ubiquity/Abundance comparison

Here we compare the method ranks for the different comparisons at alpha = 0.10.

```{r}
plotMethodRanks(c(otu_bench_file, otu_bench_file_abun, otu_bench_file_uninf), 
                colLabels = c("OTU-ubiquity","OTU-abun", "OTU-uninf"), 
                alpha = 0.10, xlab = "Comparison")
```
